---
title: "Question2"
date: "2023-12-04"
output: html_document
---

Functions used in this markdown:
```{r}

#cleaning up column names- to make them computer readable
clean_column_names <- function(penguins_data) {
  penguins_data %>%
    clean_names()
}

#Shortening species names in the data file
shorten_species <- function(penguins_data) {
  penguins_data %>%
    mutate(species = case_when(
      species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo"
    ))
}

#Removing empty columns and rows
remove_empty_columns_rows <- function(penguins_data) {
  penguins_data %>%
    remove_empty(c("rows", "cols"))
}
```


## QUESTION 2: Data Pipeline


### Introduction
I wanted to explore the relationship between body mass and flipper length in the Palmer penguins data set, and also to see if there was an interaction between these variables and the three different species. To explore these variables, I created a scatter plot- this showed a strong linear relationship between the two, as a linear line of best fit fit the data well. using this observation, I proceeded to perform an ANCOVA test based on a linear model, to see if this observed relationship was statistically significant, and fit the hypotheses I formed. Below, you can see the scatter plot I created, and the rest of my statistical method.

```{r Cleaning and exploring the data}
#Installing packages if necessary, and loading them

if (!requireNamespace("palmerpenguins", quietly = TRUE)) {
  install.packages("palmerpenguins")
}

if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("janitor", quietly = TRUE)) {
  install.packages("janitor")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}


library("ggplot2") #to plot the graphs needed
library("janitor") #to clean the data
library("palmerpenguins") #to load the palmerpenguins data
library("dplyr") #to pipe functions together

#Loading the Palmerpenguins data set, and cleaning the data
#Saving the data as a .csv file in the 'data' folder:
write.csv(penguins_raw, "data/penguins_raw.csv")

#Creating the 'penguins_clean' data- it has computer-readable column names, and any empty columns and rows are removed.
penguins_clean <- penguins_raw %>%
    clean_column_names() %>%
    shorten_species() %>%
    remove_empty_columns_rows()
#Saving this dataset as a .csv file:
write.csv(penguins_clean, "data/penguins_clean.csv")

#Creating a scatter plot to show the correlation between flipper length and body mass
FlipperLength_BodyMass_plot<- ggplot(penguins_clean, aes(x = body_mass_g, y = flipper_length_mm, colour= species)) +
  geom_point() +
    labs(title = "The relationship between body mass and flipper length",
       x = "Body Mass (g)",
       y = "Flipper Length (mm)",
       legend= "Species")+
  theme(plot.title = element_text(size = 6))+
  scale_colour_manual(values=c("#AA336A", "blue", "purple"))+
  theme_minimal()

FlipperLength_BodyMass_plot

#Installing the ragg package, to save the image
if (!requireNamespace("ragg", quietly = TRUE)) {
  install.packages("ragg")
}
library(ragg)

#Saving the objects as high resolution .png files, in the 'figures' folder

agg_png("figures/FlipperLength_BodyMass_plot.png",  width = 4000, height = 3000, units = "px", res = 600)+
print(FlipperLength_BodyMass_plot)
dev.off()
```

### Hypothesis

My hypotheses are that body mass and flipper length are positively correlated, and that there is an interaction between body mass and species. 

The null hypotheses in this test is that there is no relation or correlation between flipper length and body mass, and that there is no interaction between species and body mass, so the slopes of the graph are not significantly different from each other.


The alternative hypotheses are that there is a positive correlation, with an interaction between species and body mass, so the slopes of the graph are significantly different from each other.

### Statistical Methods

In order to test this correlation, I will create a linear model, and perform an ANCOVA test on this data. This will demonstrate whether there is a linear relationship between the two variables, while accounting for the interaction between body mass and species, and flipper length and species- if there is one.

```{r Statistics}
#Creating a linear regression model:
model <- lm(flipper_length_mm ~ body_mass_g*species, data = penguins_clean)

#Viewing the results of this model
summary(model)

#Performing an ANCOVA test
anova(model)
```

### Results & Discussion

The linear model shows that there is a statistically significant correlation between flipper length and body mass, and we can see that the p value of the intercept is <2e16. The p values for the F-ratios of both the body mass, the species, and the interaction between the two are all significant at the 95% level- indicating that there is a significant correlation between flipper length and body mass, and there is also a significant interaction between body mass and species: the body mass of each species is significantly different from each other, and the slopes of the models of each species are significantly different. We reject the null hypothesis. 

However, there are some limitations to this model- primarily, I have assumed that this data fits a linear model, in order to  perform this test. This assumption does reduce the power of this test. Also, there may be confounding variables which are affecting the results, which I have not accounted for here.

```{r Plotting Results}

# Create a results figure with ggplot2
results_figure <- ggplot(penguins_clean, aes(x = body_mass_g, y = flipper_length_mm, colour= species)) +
  geom_point() +
  geom_smooth(method=lm, linewidth=0.5)+
    labs(title = "The relationship between body mass and flipper length",
       x = "Body Mass (g)",
       y = "Flipper Length (mm)",
       legend= "Species",
       subtitle="Adjusted R-squared:  0.8553 
F-statistic: 404.2 on 5 and 336 DF,  p-value: < 2.2e-16")+
  theme(plot.title = element_text(size = 6))+
  scale_colour_manual(values=c("#AA336A", "blue", "purple"))+
  theme_minimal()

# Display the results figure
print(results_figure)

#Save the figure in the figures folder as a .png file
agg_png("figures/Results_plot.png",  width = 4000, height = 3000, units = "px", res = 600)+
print(results_figure)
dev.off()


```

In conclusion, there is a significant level of correlation between flipper length and body mass- indicating a strong, linear relationship between these two variables, even when taking into account the three different species. The results also show significant interactions between species and body mass, and that the slopes of each species are significantly different. Further research into what other elements of penguin growth correlate with flipper length would be required to be able to see whether body mass truly plays a role into determining flipper length, and if one can be estimated from the other. In the future, I would perform statistical tests on more variables to determine this.
