---
title: "Question2"
date: "2023-12-04"
output: html_document
---

Functions used in this markdown:
```{r}

#cleaning up column names- to make them computer readable
clean_column_names <- function(penguins_data) {
  penguins_data %>%
    clean_names()
}

#Shortening species names in the data file
shorten_species <- function(penguins_data) {
  penguins_data %>%
    mutate(species = case_when(
      species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie",
      species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo"
    ))
}

#Removing empty columns and rows
remove_empty_columns_rows <- function(penguins_data) {
  penguins_data %>%
    remove_empty(c("rows", "cols"))
}
```


### Introduction
I wanted to explore the relationship between body mass and flipper length in the Palmer penguins data set. To explore these variables, I created a scatter plot- this showed a strong linear relationship between the two, as a linear line of best fit fit the data well. using this observation, I proceeded to perform a linear regression test, to see if this observed relationship was staistically significant, and fit the hypotheses I formed. Below, you can see the scatter plot I created, and the rest of my statistical method.

```{r Data Exploration}
#Installing packages if necessary, and loading them

if (!requireNamespace("palmerpenguins", quietly = TRUE)) {
  install.packages("palmerpenguins")
}

if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("janitor", quietly = TRUE)) {
  install.packages("janitor")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}


library("ggplot2") #to plot the graphs needed
library("janitor") #to clean the data
library("palmerpenguins") #to load the palmerpenguins data
library("dplyr") #to pipe functions together

#Loading the Palmerpenguins data set, and cleaning the data
#Saving the data as a .csv file in the 'data' folder:
write.csv(penguins_raw, "data/penguins_raw.csv")

#Creating the 'penguins_clean' data- it has computer-readable column names, and any empty columns and rows are removed.
penguins_clean <- penguins_raw %>%
    clean_column_names() %>%
    shorten_species() %>%
    remove_empty_columns_rows()
#Saving this dataset as a .csv file:
write.csv(penguins_clean, "data/penguins_clean.csv")

#Creating a scatter plot to show the correlation between flipper length and body mass
FlipperLength_BodyMass_plot<- ggplot(penguins_clean, aes(x = body_mass_g, y = flipper_length_mm, colour= species)) +
  geom_point() +
    labs(title = "The relationship between body mass and flipper length",
       x = "Body Mass (g)",
       y = "Flipper Length (mm)",
       legend= "Species")+
  theme(plot.title = element_text(size = 6))+
  scale_colour_manual(values=c("#AA336A", "blue", "purple"))+
  theme_minimal()

FlipperLength_BodyMass_plot
#Installing the ragg package, to save the image
if (!requireNamespace("ragg", quietly = TRUE)) {
  install.packages("ragg")
}
library(ragg)

#Saving the objects as high resolution .png files, in the 'figures' folder

agg_png("figures/FlipperLength_BodyMass_plot.png",  width = 4000, height = 3000, units = "px", res = 600)+
print(FlipperLength_BodyMass_plot)
dev.off()
```

### Hypothesis

My hypothesis is that body mass and flipper length are positively correlated, and that there is an interaction between body mass and species. The null hypothesis in this test is that there is no relation or correlation between flipper length and body mass, and that there is no interaction between species and body mass, and the alternative hypothesis is that there is a positive correlation, with an interaction between species and body mass.

### Statistical Methods

In order to test this correlation, I will create a linear model, and perform a linear regression test on this data. This will demonstrate whether there is a linear relationship between the two variables, while accounting for the interaction between body mass and species, and flipper length and species- if there is one.

```{r Statistics}
#Creating a linear regression model:
model <- lm(flipper_length_mm ~ body_mass_g*species, data = penguins_clean)

#Viewing the results of this model
summary(model)
```

### Results & Discussion
The p value of the coefficient of body_mass_g is very low, indicating a high level of statistical significance- even after accounting for species, there is a strong positive correlation between body mass and flipper length. The p value of the coefficient for the two species are not significant, however, indicating that there is not a significant difference in the flipper length of the three species. There is a significant p value in the interactions between body mass and the species, however- indicating that the different penguin species have significantly different body masses. The adjusted R-squared is 0.8553, showing that 85.53% of the variation in flipper length is explained by the model- this is statistically significant according to the very low p value of less than 2.2e-16. This correlation does not imply causation, however- it is possible that this relationship does not mean anything in terms of one variable influencing the other. Looking at the two, however, it seems feasible to say that individuals with higher body mass are probably larger overall, so they are likely to have bigger body parts, including longer flippers.

```{r Plotting Results}

# Create a results figure with ggplot2
results_figure <- ggplot(penguins_clean, aes(x = body_mass_g, y = flipper_length_mm, colour= species)) +
  geom_point() +
  geom_smooth(method=lm, linewidth=0.5)+
    labs(title = "The relationship between body mass and flipper length",
       x = "Body Mass (g)",
       y = "Flipper Length (mm)",
       legend= "Species",
       subtitle="Adjusted R-squared:  0.8553 
F-statistic: 404.2 on 5 and 336 DF,  p-value: < 2.2e-16")+
  theme(plot.title = element_text(size = 6))+
  scale_colour_manual(values=c("#AA336A", "blue", "purple"))+
  theme_minimal()

# Display the results figure
print(results_figure)

#Save the figure in the figures folder as a .png file
agg_png("figures/Results_plot.png",  width = 4000, height = 3000, units = "px", res = 600)+
print(results_figure)
dev.off()


```

In conclusion, there is a significant level of correlation between flipper length and body mass- indicating a strong, linear relationship between these two variables, even when taking into account the three different species. The results show significant interactions between species and body mass, but not between species and flipper length, despite the strong correlation between body mass and flipper length. Further research into what other elements of penguin growth correlate with flipper length would be required to be able to see whether body mass truly plays a role into determining flipper length, and if one can be estimated from the other. In the future, I would perform statistical tests on more variables to determine this.
